---
title: "Analysis of weather station data using hoboR"
output: html_document
---

## Installation and Setup


```{r installation, include=TRUE, message=FALSE, warning=FALSE}
#install.packages("devtools", quietly = TRUE)
#install.packages("reshape", quietly = TRUE)

# Load libraries 
library(reshape)
library(lubridate)
library(tidyr)
library(ggplot2)
library(hoboR)

```

## Site 12
The data was collected in China Creek in Southern Oregon using a weather station. The measurements were recorded every minute over a period of 5 months, from in Summer and Fall 2022. The weather variables collected were humidity (Wetness), temperature (Temp), relative humidity (RH), and rain (Rain). 

```{r addpath, include=FALSE}
# Add the PATH to your sites for weather data (from hobo)
path = paste0("~/Desktop/Adam/site_12_date_adj2/")
# make sure the path to your CSV files exists
file.exists(path)     # this will return a logical value TRUE
```
Confirm that the path exist, then proceed with binding all the csv files, and clean the data.
```{r cleaning, include=T}
# loading all hobo files
hobofiles <- hobinder(path, skip = 1)
hobocleaned <- hobocleaner(hobofiles, format = "yymd")
head(hobocleaned)
```

Let's summarize the data every 30 min, and get the means for 1 day or get the mean every 24 hours

> Note: that the original data was recorded every minute.

```{r summary, include=T}
# getting hobo mean summary by time
hobot5 <- hobotime(hobocleaned, summariseby = "30 mins", na.rm = T)
hobomeans5 <- meanhobo(hobot5, summariseby = "1 day",  na.rm = T)
head(hobomeans5)

# getting hobo means by date
hobomeans <- meanhobo(hobocleaned, summariseby = "24 h",  na.rm = T)
head(hobomeans)
```
Check the difference between both methods, summarizing and getting the mean, or mean only.
```{r plot, include=TRUE, echo= TRUE}
plot(1:nrow(hobomeans), hobomeans$x.Temp, type = "line")
lines(1:nrow(hobomeans5), hobomeans5$x.Temp, type = "line", col = "red")
```

If you are interested to spot a time range, you can specify the dates using `hoborange`, and aadd the starting and end dates of interest.
```{r specify, include=TRUE}
# Specify a window range 
timerange <- hoborange(hobocleaned, start="2022-08-08", end="2022-12-12")
head(timerange)
```
Let's check the variability every 24 hours at midnight and noon for 100 days. Store it in a variable, and add groups. 

```{r snapshot, include=TRUE, echo= TRUE}
# Snapshot of a time interval 
a <- timestamp(hobocleaned, stamp = "2022-08-05 00:01", by = "24 hours",
          days = 100, na.rm = TRUE, plot = F, var = "Temp")
a$Group <- rep("night", nrow(a))
b <- timestamp(hobocleaned, stamp = "2022-08-05 12:01", by = "24 hours",
               days = 100, na.rm = TRUE, plot = F, var = "Temp")
b$Group <- rep("day", nrow(b))

daynight <- rbind(a, b)
```

Now, you can plot with `ggplot` library 

```{r ggplot, fig.cap="Temperature Variability: Day vs. Night"}
ggplot(daynight, aes(x = as.POSIXct(Date), y = Temp, group = Group, color = Group)) +
  geom_line() +
  scale_x_datetime() +
  xlab(NULL)+
  scale_y_continuous(limits = c(0, 30)) +
  scale_color_manual(values = c("orange", "black")) +
  labs(color = "Source") +
  theme_minimal()
```

```{r ggplot2, fig.cap="Two Variables: Temp vs. Humidity"}
ggplot(hobocleaned, aes(x=as.POSIXct(Date))) +
  geom_line( aes(y=Temp, col = "Temp"), alpha = 0.5) + 
  geom_line( aes(y= Wetness, col = "Wetness"), alpha = 0.5) + 
  scale_y_continuous(
    # Features of the first axis
    name = "Temperature °C",
    # Add a second axis and specify its features
    # limits = c(min(hobocleaned$Temp-5, na.rm = TRUE), max(hobocleaned$Temp+5, na.rm = TRUE)), # Limits for Temp
    sec.axis = sec_axis(~ ., name = "Wetness") # Secondary axis for Wetness
  ) +
  labs(title = "Temperature: Oct 14 - Nov 11, 2021", color = "Legend") +
  scale_color_manual(labels = c("Humidity", "Temp"), values = c("blue", "red")) +
  scale_x_datetime(labels = date_format("%Y-%m-%d"))+
  theme_bw()
```
